TG2 Publications

### 2025-08-11 activities.qmd
Revision 250707, small amendment
Minor updates by Aug 11, 2025 (GH)

* Invited session at 70th Annual Meeting of the German Society for Medical Informatics, Biometrics and Epidemiology (Deutsche Gesellschaft für Medizinische Informatik, Biometrie und Epidemiologie e. V., GMDS) in Jena, Germany, 7-11 Sep 2025. Speaker for TG2: Gregor Buch. Topic of the talk: STRATOS Topic Group 2 – Selection of variables and functional forms in multivariable analysis: Results from a review of Covid-19 prognostic models.[Find out more (in German)](https://gmds2025.de/)

* Minisymposium at the International Society of Clinical Biostatistics (ISCB) conference in Basel, CH, 28 Aug 2025. Topic of the session: Statistical Research needs to improve – on the important roles of simulation studies and guidance for analysis. Speakers (among others): Georg Heinze, Willi Sauerbrei, Michal Abrahamowicz [Find out more](https://iscb2025.info/mini-symposia.html#mini-symposium-3-statistical-research-needs-to-improve-on-the-important-roles-of-simulation-studies-and-guidance-for-analysis)

@@

* Invited talk at the [ROeS 2025](https://roes2025.at/index.html) (Conference of the Austro-Swiss Region of the International Biometric Society) in Graz, A, 14-18 Sep 2025. Topic of the session: Towards best practices in medical statistics: The need to implement innovative scientific solutions. Speaker: Theresa Ullmann. Topic of the talk: Neutral Comparisons Of Statistical Methods: Opportunities, Challenges, And Common Issues

>>>>

* Minisymposium at the International Society of Clinical Biostatistics (ISCB) conference in Basel, CH, 28 Aug 2025. Topic of the session: Statistical Research needs to improve – on the important roles of simulation studies and guidance for analysis. Speakers (among others): Georg Heinze, Willi Sauerbrei, Michal Abrahamowicz. [Find out more](https://iscb2025.info/mini-symposia.html#mini-symposium-3-statistical-research-needs-to-improve-on-the-important-roles-of-simulation-studies-and-guidance-for-analysis)

* Invited session at 70th Annual Meeting of the German Society for Medical Informatics, Biometrics and Epidemiology (Deutsche Gesellschaft für Medizinische Informatik, Biometrie und Epidemiologie e. V., GMDS) in Jena, Germany, 7-11 Sep 2025. Speaker for TG2: Gregor Buch. Topic of the talk: STRATOS Topic Group 2 – Selection of variables and functional forms in multivariable analysis: Results from a review of Covid-19 prognostic models. [Find out more (in German)](https://gmds2025.de/)

@@

* Invited talk at the [ROeS 2025](https://roes2025.at/index.html) (Conference of the Austro-Swiss Region of the International Biometric Society) in Graz, A, 14-18 Sep 2025. Topic of the session: Towards best practices in medical statistics: The need to implement innovative scientific solutions. Speaker: Theresa Ullmann. Topic of the talk: Neutral Comparisons Of Statistical Methods: Opportunities, Challenges, And Common Issues (Includes as an example TG2-P11, Simulation study on variable selection)

### 2025-08-11 index.qmd
Minor updates by Aug 11, 2025 (GH)

![2nd General Meeting of the STRATOS members at the Banff International Research Station (BIRS), Banff, Canada.](assets/STRATOS_Banff_second_meeting.png){fig-align="center"}

>>>>

![2nd General Meeting of the STRATOS members at the Banff International Research Station (BIRS), Banff, Canada, June 2019.](assets/STRATOS_Banff_second_meeting.png){fig-align="center"}

### 2025-08-11 members.qmd
Corrected affiliation
added CSL
Minor updates by Aug 11, 2025 (GH)

**Aris Perperoglou (chair)**:  Data Science & AI,   AstraZeneca R&D Biopharmaceuticals , Cambridge UK

**Willi Sauerbrei (chair)**: Institute for Medical Biometry and Medical Informatics, Medical Center - University of Freiburg, Germany

@@

**Daniela Dunkler**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria

@@

**Mariana Nold**: Department of Sociology, Friedrich-Schiller University of Jena, Jena, Germany

@@

@@

**Theresa Ullmann**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria

>>>>

**[Aris Perperoglou](https://uk.linkedin.com/in/arisperperoglou) (chair)**:  Predictive Modelling, GSK, Stevenage, UK

**[Willi Sauerbrei](https://www.uniklinik-freiburg.de/english/imbi/mitarbeiter.html?imbiuser=wfs) (chair)**: Institute for Medical Biometry and Medical Informatics, Medical Center - University of Freiburg, Germany

@@

**[Daniela Dunkler](https://www.meduniwien.ac.at/researcher/daniela_dunkler)**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria

@@

**[Mariana Nold](https://www.fsv.uni-jena.de/19575/dr-mariana-nold)**: Department of Sociology, Friedrich-Schiller University of Jena, Jena, Germany

@@

## Clinical affiliate

**Carl-Stephan Leonhardt**: Department of Surgery, Medical University of Vienna, Vienna, Austria

@@

**[Theresa Ullmann](https://www.meduniwien.ac.at/researcher/theresa_ullmann)**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria

### 2025-08-11 papers.qmd
Test update on papers.qmd
Full update
Minor updates by Aug 11, 2025 (GH)

@@

Aris Perperoglou , Georg Heinze , Willi Sauerbrei on behalf of STRATOS TG2

[Link](https://www.stratos-initiative.org/sites/default/files/2021-11/STRATOS-TG2-BB-variable-selection.pdf)

@@

# Sugggested Publications

@@

@@

* Heinze G, Wallisch C, Dunkler D. Variable selection – a review and recommendations for the practicing statistician. Biometrical J. 2018; 60:431–49.

* Binder H, Sauerbrei W, Royston P. Comparison between splines and fractional polynomials for multivariable model building with continuous covariates: a simulation study with continuous response. Statistics in Medicine 2013; 32(13): 2262– 2277.

>>>>

## Recent activities of the Topic Group on Selection of Variables and Functional Forms in Multivariable Analysis (TG2)

Georg Heinze, Aris Perperoglou, Willi Sauerbrei on behalf of STRATOS TG2

[Biometric Bulletin 2021; 38(2):7-8](https://www.stratos-initiative.org/sites/default/files/2021-09/IBS_BioMetric-Bulletin-Vol38-Issue2-1_4-7-8.pdf)

<details>
<summary>Abstract</summary>
This Biometric Bulletin article provides an update of the activities of TG2 by July 2021.
</details>

@@

Aris Perperoglou, Georg Heinze, Willi Sauerbrei on behalf of STRATOS TG2

[Biometric Bulletin 2018; 35(3):18-19](https://www.stratos-initiative.org/sites/default/files/2021-11/STRATOS-TG2-BB-variable-selection.pdf)

@@

# Suggested Publications

@@

* Kipruto, E., & Sauerbrei, W. (2025). Evaluating Prediction Performance: A Simulation Study Comparing Penalized and Classical Variable Selection Methods in Low-Dimensional Data. Applied Sciences, 15(13), 7443. [https://doi.org/10.3390/app15137443](https://doi.org/10.3390/app15137443)

* Sauerbrei, W., Royston, P., & Kipruto, E. (2025). Multivariable Fractional Polynomial Models and Extensions. In International Encyclopedia of Statistical Science (pp. 1609-1616). Berlin, Heidelberg: Springer. [https://doi.org/10.1007/978-3-662-69359-9_397](https://doi.org/10.1007/978-3-662-69359-9_397)

* Kipruto, E., & Sauerbrei, W. (2024). Post‐Estimation Shrinkage in Full and Selected Linear Regression Models in Low‐Dimensional Data Revisited. Biometrical Journal, 66(7), e202300368. [https://doi.org/10.1002/bimj.202300368](https://doi.org/10.1002/bimj.202300368)

@@

* Heinze G, Wallisch C, Dunkler D. Variable selection – a review and recommendations for the practicing statistician. Biometrical J. 2018; 60:431–49. 
* Binder H, Sauerbrei W, Royston P. Comparison between splines and fractional polynomials for multivariable model building with continuous covariates: a simulation study with continuous response. Statistics in Medicine 2013; 32(13): 2262– 2277.
